---
title: "Activity Recognition"
author: "Juan A. Reyes D."
date: "February 20th 2018"
---

# Data Source

```{r}
TrainLink<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TestLink  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

# Reading Data

```{r}
TrainFile<-"pml-traininig.csv"
TestFile<-"pml-testing.csv"

download.file(TrainLink,destfile = TrainFile)
download.file(TestLink, destfile = TestFile)

TrainData <- read.csv(TrainFile)
TestData  <- read.csv(TestFile)
```

# Partitioning and Cleaning Dataset

The training dataset is partinioned in 2 to create a Train data (70% of the data) for the modeling process and a Test Data (with the remaining 30%) for the validations. 

```{r}
library(caret)

PartitionTrain  <- createDataPartition(TrainData$classe, p=0.7, list=FALSE)
Train <- TrainData[PartitionTrain, ]
Test  <- TrainData[-PartitionTrain, ]

#remove variables with Nearly Zero Variance
NearZeroVar <- nearZeroVar(Train)
Train <- Train[, -NearZeroVar]
Test  <- Test[, -NearZeroVar]
dim(TestSet)

#remove variables that are mostly NA
FullNA    <- sapply(Train, function(x) mean(is.na(x))) > 0.95
Train <- Train[, FullNA==FALSE]
Test  <- Test[, FullNA==FALSE]

#remove identification only variables (columns 1 to 5)
Train <- Train[, -(1:5)]
Test  <- Test[, -(1:5)]
```

# Prediction Model Building with Random Forest

```{r}
library(randomForest)
library(rpart)
library(rpart.plot)

set.seed(666)
RF <- trainControl(method="cv", number=3, verboseIter=FALSE)
FitRandForest <- train(classe ~ ., data=Train, method="rf", trControl=RF)
FitRandForest$finalModel
```

# Prediction on Test dataset

```{r}
predictRF <- predict(FitRandForest, newdata=Test)
confMatRF <- confusionMatrix(predictRF, Test$classe)
confMatRF

#plot matrix results
plot(confMatRF$table, col = confMatRF$byClass, main = paste("Random Forest - Accuracy =",
     round(confMatRF$overall['Accuracy'], 4)))
```     
     
# Conclusion

As can be seen from the confusion matrix the Random Forest model is very accurate, about 99%. And all the 20 answers were validated as correct at the PML.
